---
title: "team"
author: "Adam Ford and Kellan Bouwman"
date: "9/24/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Setup
```{r}
# dplyr
library(dplyr)
# ggplot2
library(ggplot2)



dat <- readxl::read_xls('GSS(2).xls')

dat <- rename(dat, marriage = Marital_status, happiness = General_happiness, polParty = Political_party_affiliation, avgWorkWeek = Number_of_hours_usually_work_a_week, highEd = Highest_year_of_school_completed, gender = Respondents_sex)
```

## 1
```{r}
dat$highEd <- as.numeric(as.character(dat$highEd))
dat %>%
  group_by(polParty) %>%
  summarise_at(vars(highEd), funs(mean(., na.rm=TRUE))) -> grouped

ndx = order(grouped$highEd, decreasing = TRUE)

grouped[ndx,]

```
     
    Answer: Generally republicans have the higher average years of education than democrats, whether that be an identified republican or near. Additionally, Those who are unsure or dont answer have the least education, this may just be because of age.

## 2
```{r}
filtered <- filter(dat, (polParty == 'Don\'t know' & highEd >= 11) | (polParty == 'Ind,near dem' & highEd >= 13) | ( polParty == 'Ind,near rep' & highEd >= 13) | (polParty == 'Independent' & highEd >= 12) | (polParty == 'No answer' & highEd >= 12) | (polParty == 'Not str democrat' & highEd >= 12) | ( polParty == 'Not str republican' & highEd >= 13) | (polParty == 'Other party' & highEd >= 13) |  (polParty == 'Strong democrat' & highEd >= 12) | (polParty == 'Strong republican' & highEd>= 12))
filtered
summary(filtered)
str(filtered)
```

    Answer: After filtering the data set contains the same columns, but with about 2/3 of the rows. This is due to allowing those with an education equal to the median to be included. The easiest way to be sure the dataset is filtered is to look at the summary of education, where it shows the minimum is 11, which is the lowest median of any political party. 

## 3
```{r}

marriageData <- dat %>%
  select(happiness, marriage, polParty) %>%  mutate(recode(happiness, "Very happy" = "Happy", "Pretty happy" = "Happy"))
marriageData$happiness = recode(marriageData$happiness, "Very happy" = "Happy", "Pretty happy" = "Happy", "Not too happy" = "Not Happy", "Don't know" = "NA", "Not applicable" = "NA", "No answer" = "NA")

marriageData %>%
  group_by(marriage, happiness) %>%
  summarize(count = n())

ggplot(data = marriageData, aes(x = marriage, fill = marriage)) + geom_bar(width=0.7, stat='count') + facet_wrap(~happiness, drop = TRUE, scales = "free_y") + theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank())
```

    Answer: Yes because in the happy catergory people who are married tend to be happier signifigacntly at a higher rate. If you look at the amount of people in happy and compare the those who are married to not married, people who are married are double that of never married which is double that of divorded people who are happy. Intrestling divored and widowed happiness are about similar. In not happ the overall population is signifiagntly smaller, but the not happy count between married and never married is about the same, where as the divorced and widowed which are simlar are apporximatly 30% less, with the lowest count being seperated people.  

## 4
```{r}
marriageData %>%
  group_by(marriage, happiness, polParty) %>%
  summarize(count = n())

ggplot(data = marriageData, aes(x = polParty, fill = polParty)) + geom_bar() + facet_grid(happiness ~ marriage, drop = TRUE, scales = "free_y") + theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank())
```

    Answer: There is not much difference in the data, instead there is large amount of similarities looking at a particular intercetion the largest populations are not strong democrat or republic which show that is has little or no affect on happiness. 
    
## 5
```{r}
dat %>%
  group_by(gender, marriage) %>%
  summarize(count = n())


ggplot(data = dat, aes(x = marriage, fill = marriage)) + geom_bar(width=0.7) + facet_wrap(~gender, drop = TRUE) + theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank())
```

    Answer: Women of all marital status execpt never married had a higher response count. Even in the never married case they are fairly close. This shows that more women respond to the survey then men do. This could just be representative that there are more women in the survey or that men chose to not answer the survey. 

## 6 Part 1
```{r}
dat %>%
  group_by(gender) %>%
  summarise_at(vars(highEd), funs(mean(., na.rm=TRUE))) -> groupedGender
groupedGender

ggplot(dat, aes(x=gender, y=highEd)) + geom_boxplot()
```

    Answer: Here we can see the relationship between the respondents sex and the level of education they achieved. Numerically we can see that there is some difference, that men average 13 years and females average 12.8 years. The boxplot shows a similar story that they are very similar. They both hold the same median and Q1 boundary at 12. We do see a difference in the Q3 boundary where extends to 16 and women stop at 15. This boxplot may not be the perfect representation as the scale isn't completely continuous. 


## 6 Part 2
```{r}
dat$avgWorkWeek <- as.numeric(as.character(dat$avgWorkWeek))
dat %>%
  group_by(happiness) %>%
  summarise_at(vars(avgWorkWeek), funs(mean(., na.rm=TRUE))) -> workWeekByHappy

workWeekByHappy

ggplot(dat, aes(x=happiness, y=avgWorkWeek)) + geom_boxplot()
```

    Answer: While the numerical summary shows that the average hours worked is fairly close no matter how happy you are with 38.3 being the mean for not applicable, not too happy, and pretty happy. Now the intresting thing is that based soley off the numerical summary the very happy people actually worked more at 40.2 hours. This is also where the box plot comes into play, we can see that the pretty happy and very happy have large amount of outliers and thus a bit more varity to their work schedule. Also there appears to be more people in the pretty happy and very happy so there is mostly other factors that also affect happiness and the hours you work on average.  
